---
title: "NHL Data Analysis"
author: "Andrew Daly"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    fig_caption: true
    css: "custom-style.css"
params:
  state: "hawaii"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisites

You will need GGPlot and XML2.
These can be insalled by running:  
  **install.packages("ggplot2")**  
  **install.packages("xml2")**  
  
  
## Details
In this assignment I will demonstrate how to scrape, clean and combine data sets. In addition, I will demonstrate how to generate summary statistics and simple plots. I will be scraping initial NHL data from [Scrape This Site](https://www.scrapethissite.com/pages/forms/).  


Clear your environment.
```{r}
rm(list = ls())
```
  
  
### Create initial data frame with scraped data

Load XML library to parse results from the web pages.
```{r}
library(xml2)
```
  
  
Initialize empty vectors to hold data scraped from the web pages.
```{r}
all_team <- character(0)
all_year <- character(0)
all_wins <- character(0)
all_losses <- character(0)
```
  
  
Scrape the hockey data from _Scrape This Site_
```{r}
for (i in 1:24) { # 24 pages of data

  Sys.sleep(5) # Add 5-second wait time

  # Paste number (1, 2, 3, ...) at end of base URL
  url <- paste("https://www.scrapethissite.com/pages/forms/?page_num=", i, sep = "")
  page <- read_html(url)

  team <- xml_text(xml_find_all(page, "//tr/td[1]"))
  all_team <- c(all_team, team) # Append to names vector

  year <- xml_text(xml_find_all(page, "//tr/td[2]"))
  all_year <- c(all_year, year) # Append to years vector

  wins <- xml_text(xml_find_all(page, "//tr/td[3]"))
  all_wins <- c(all_wins, wins) # Append to wins vector

  losses <- xml_text(xml_find_all(page, "//tr/td[4]"))
  all_losses <- c(all_losses, losses) # Append to losses vector

  # Wait a random amount of time before moving to the next page
  random_value <- sample(2:20, 1)
  Sys.sleep(random_value)
}
```
  
  
  
Combine individual vectors into a data frame.
```{r}
nhl0 <- data.frame(all_team, all_year, all_wins, all_losses)
colnames(nhl0) <- c("team", "year", "wins", "losses")
```


### Clean the data.  

 - Remove white space.
```{r}
nhl0$team <- trimws(nhl0$team)
nhl0$year <- trimws(nhl0$year)
nhl0$wins <- trimws(nhl0$wins)
nhl0$losses <- trimws(nhl0$losses)
```
  
  
Convert Year, Wins, and Losses columns to Integer data type.
```{r}
nhl0$year <- as.integer(nhl0$year)
nhl0$wins <- as.integer(nhl0$wins)
nhl0$losses <- as.integer(nhl0$losses)
```
  
  
Review the column names and data types in the data frame.
```{r}
str(nhl0)
```
  
  
Review the first 10 rows of data in the data frame.
```{r echo=FALSE}
nhl0[1:10, ]
```
  

### Import additional data for integration

Read excel file for more corresponding NHL data.
```{r}
library(readxl)
nhl_expanded <- read_excel("nhl_2012-2021.xlsx")
```  


- Clean and transform 2012-2021 NHL data to integrate with scrapped data.
```{r echo=FALSE}
names(nhl_expanded)[names(nhl_expanded) == "Team"] <- "team"
names(nhl_expanded)[names(nhl_expanded) == "W"] <- "wins"
names(nhl_expanded)[names(nhl_expanded) == "L"] <- "losses"
names(nhl_expanded)[names(nhl_expanded) == "Season"] <- "year"
nhl_expanded$year <- as.integer(nhl_expanded$year)
nhl_expanded$wins <- as.integer(nhl_expanded$wins)
nhl_expanded$losses <- as.integer(nhl_expanded$losses)

nhl_expanded$team <- gsub("\\*$", "", nhl_expanded$team)
```  


Remove additional columns in NHL expanded data frame by creating another data frame with relevant columns.
```{r}
nhl_expanded2 <- data.frame(nhl_expanded$team, nhl_expanded$year, nhl_expanded$wins, nhl_expanded$losses)
colnames(nhl_expanded2) <- c("team", "year", "wins", "losses")
```


Merge the two data frames vertically. The data will share the same variables and increase in observation depth
```{r}
nhl2 <- rbind(nhl0, nhl_expanded2)
```


Create a new variable for win percentage. The table below shows the addition of the win percentage column.
``` {r echo=FALSE}
nhl2$Win_percentage <- round((nhl2$wins / (nhl2$wins + nhl2$losses)), 2)
```
```{r echo=FALSE}
data <- head(nhl2, 4)
knitr::kable(data)
```


### Import NHL arena data for integration


Import and read the NHL arenas comma separated values file.
```{r}
arena <- read.csv("nhl_hockey_arenas.csv")
```


Rename the columns in arena to match the other data frame for integration.
```{r echo=FALSE}
names(arena)[names(arena) == "Team.Name"] <- "team"
names(arena)[names(arena) == "Arena.Name"] <- "arena"
names(arena)[names(arena) == "Arena.Location"] <- "location"
names(arena)[names(arena) == "Seating.Capacity"] <- "capacity"
```


Drop unnecessary columns in the arena data frame.
```{r}
arena$Opening.Year <- NULL
```


Output the first 5 rows of data in the arena data frame.
```{r echo=FALSE}
head(arena, 5)
```


Check for spelling errors within the team name column. Some teams may have adapted new names over the years.
```{r echo=FALSE}
error <- unique(arena$team)
error
```


Correct the misspelled data and modernize team names from the same franchise.
```{r}
arena$team <- ifelse(arena$team == "Seattle Kracken", "Seattle Kraken", arena$team)
nhl2$team <- ifelse(nhl2$team == "Mighty Ducks of Anaheim", "Anaheim Ducks", nhl2$team)
```


Merge the data horizontally, whereby adding more columns to the existing data frame
```{r}
nhl <- merge(nhl2, arena, by = "team", all.x = TRUE)
head(nhl, 5)
```


The New York Islanders have two accounts of an arena. They built a new arena in 2012, write if else statements to correctly match the year to the arena for the New York Islanders.
```{r}
nhl$arena <- ifelse(nhl$team == "New York Islanders",
  ifelse(nhl$year >= 2012, "Nassau Coliseum",
    ifelse(nhl$year < 2012, "Barclays Center", nhl$arena)
  ),
  nhl$arena
)

nhl$location <- ifelse(nhl$team == "New York Islanders",
  ifelse(nhl$arena == "Nassau Coliseum", "Uniondale, New York",
    ifelse(nhl$arena == "Barclays Center", "Brooklyn, New York", nhl$location)
  ),
  nhl$location
)

nhl$capacity <- ifelse(nhl$team == "New York Islanders",
  ifelse(nhl$arena == "Nassau Coliseum", 13900,
    ifelse(nhl$arena == "Barclays Center", 15795, nhl$capacity)
  ),
  nhl$capacity
)
```


The above command creates exact duplicates, multiple rows containing the same data. Remove those rows.
```{r}
nhl <- nhl[!duplicated(nhl), ]
```


Save the final data frame to a comma separated values file without row names.
```{r}
write.csv(nhl, "hockey_data.csv", row.names = FALSE)
```


## Data Modeling 

### 1990 Teams Winning More Than 50% Of Regular Season Games

Embed ggplot listing all NHL teams that won more than 40 of the 80 regular season games in 1990.
```{r pressure, echo=FALSE}
library(ggplot2)

ggplot(nhl[which(nhl$year == 1990 & nhl$wins > 40), c("team", "wins")], aes(x = reorder(team, wins), y = wins, fill = team)) +
  geom_bar(stat = "identity", width = .75) +
  coord_flip() +
  theme(legend.position = "none") +
  ggtitle("NHL Teams > 40 Wins in 1990") +
  xlab("NHL Team") +
  ylab("Wins")
```
  

### Philadelphia Flyers Wins Each Season
```{r pressure1, echo=FALSE}
flyers_data <- nhl[nhl$team == "Philadelphia Flyers", ]
ggplot(flyers_data, aes(x = year, y = wins, color = team)) +
  geom_line() +
  ggtitle("Philadelphia Flyers Wins Over the Years") +
  xlab("Year") +
  ylab("Wins") +
  theme_minimal()
```


### Distribution of Win Percentages by Team
```{r pressure2, echo=FALSE}
ggplot(nhl, aes(x = team, y = Win_percentage, fill = team)) +
  geom_boxplot() +
  ggtitle("Distribution of Win Percentages by Team") +
  xlab("NHL Team") +
  ylab("Win Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### NHL Win Percentage Over the Years
```{r pressure3, echo=FALSE}
ggplot(nhl, aes(x = year, y = Win_percentage, fill = team)) +
  geom_bar(stat = "identity") +
  ggtitle("NHL Win Percentage Over the Years") +
  xlab("Year") +
  ylab("Win Percentage") +
  theme_minimal()
```


### NHL Arena Capacities
```{r pressure4, echo=FALSE}
ggplot(nhl, aes(x = reorder(team, -capacity), y = capacity, fill = team)) +
  geom_bar(stat = "identity") +
  ggtitle("NHL Arena Capacities") +
  xlab("NHL Team") +
  ylab("Capacity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)
```

  